---
title: "p8105_hw6_xw2972"
author: "Xiwen Wu"
output: github_document
date: 2024-12-02
---
##Question 1:
##The 95% confidence interval for R-squared [0.89, 0.93] shows the range of likely values for this statistic, reflecting the variability due to sampling.
##The 95% confidence interval for beta product [1.96, 2.06] shows the plausible range of this derived parameter, offering insight into its stability.
```{r}
library(rnoaa)
library(dplyr)
library(ggplot2)
library(broom)

weather_df <- rnoaa::meteo_pull_monitors(
  c("USW00094728"),
  var = c("PRCP", "TMIN", "TMAX"), 
  date_min = "2017-01-01",
  date_max = "2017-12-31"
) %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10
  ) %>%
  select(name, id, everything())

n_bootstrap <- 5000

set.seed(123) 
bootstrap_results <- replicate(n_bootstrap, {

  sampled_data <- weather_df %>% sample_frac(size = 1, replace = TRUE)
  
  lm_fit <- lm(tmax ~ tmin, data = sampled_data)
  
  r_squared <- glance(lm_fit)$r.squared
  
  coefs <- tidy(lm_fit) %>% pull(estimate)
  log_beta_product <- log(coefs[1] * coefs[2])
  
  c(r_squared, log_beta_product)
}, simplify = TRUE)


bootstrap_df <- as.data.frame(t(bootstrap_results))
names(bootstrap_df) <- c("r_squared", "log_beta_product")

ci_r_squared <- quantile(bootstrap_df$r_squared, probs = c(0.025, 0.975))
ci_log_beta_product <- quantile(bootstrap_df$log_beta_product, probs = c(0.025, 0.975))

ggplot(bootstrap_df, aes(x = r_squared)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Bootstrap Distribution of R-Squared",
       x = "R-Squared", y = "Density") +
  theme_minimal()

ggplot(bootstrap_df, aes(x = log_beta_product)) +
  geom_density(fill = "green", alpha = 0.5) +
  labs(title = "Bootstrap Distribution of log(β0 * β1)",
       x = "log(β0 * β1)", y = "Density") +
  theme_minimal()

list(
  r_squared_CI = ci_r_squared,
  log_beta_product_CI = ci_log_beta_product
)
```
##Question 2:
##Fresno CA has the highest adjusted ratio, while New York NY has the lowest adjusted ratio.
```{r}
library(dplyr)
library(broom)
library(ggplot2)
library(purrr)
library(tidyr)

homicide_data <- read.csv("homicide-data.csv")

homicide_data <- homicide_data %>%
  mutate(
    city_state = paste(city, state, sep = ", "),
    solved = ifelse(disposition %in% c("Closed by arrest", "Closed without arrest"), 1, 0)
  ) %>%
  filter(
    !(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")),
    victim_race %in% c("White", "Black"),
    !is.na(victim_age)
  ) %>%
  mutate(victim_age = as.numeric(victim_age))

baltimore_data <- homicide_data %>%
  filter(city_state == "Baltimore, MD")

baltimore_glm <- glm(
  solved ~ victim_age + victim_sex + victim_race,
  data = baltimore_data,
  family = binomial
)

baltimore_summary <- broom::tidy(baltimore_glm, conf.int = TRUE) %>%
  filter(term == "victim_sexMale") %>%
  mutate(
    adjusted_or = exp(estimate),
    ci_lower = exp(conf.low),
    ci_upper = exp(conf.high)
  )

baltimore_summary

city_results <- homicide_data %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~ glm(solved ~ victim_age + victim_sex + victim_race, 
                            data = ., family = binomial)),
    tidy_model = map(model, ~ broom::tidy(., conf.int = TRUE)),
    or_data = map(tidy_model, ~ .x %>% 
                    filter(term == "victim_sexMale") %>%
                    mutate(
                      adjusted_or = exp(estimate),
                      ci_lower = exp(conf.low),
                      ci_upper = exp(conf.high)
                    ))
  ) %>%
  unnest(or_data, keep_empty = TRUE) %>%
  select(city_state, adjusted_or, ci_lower, ci_upper)

city_results <- city_results %>%
  arrange(adjusted_or)

ggplot(city_results, aes(x = reorder(city_state, adjusted_or), y = adjusted_or)) +
  geom_point() +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +
  coord_flip() +
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides by City",
    x = "City, State",
    y = "Adjusted Odds Ratio (Male vs Female Victims)"
  ) +
  theme_minimal()
```
##Question 3:
```{r}
library(rsample)
library(modelr)
library(broom)
library(ggplot2)
library(purrr)

birthweight_df <- read.csv("birthweight.csv")

birthweight_df$babysex <- as.factor(birthweight_df$babysex)
birthweight_df$frace <- factor(birthweight_df$frace, levels = c(1, 2, 3, 4, 8, 9))
birthweight_df$mrace <- factor(birthweight_df$mrace, levels = c(1, 2, 3, 4, 8))

model_1 <- lm(bwt ~ gaweeks + delwt + babysex, data = birthweight_df)

birthweight_df <- add_residuals(birthweight_df, model_1, var = "resid_1")
birthweight_df <- add_predictions(birthweight_df, model_1, var = "fitted_1")

ggplot(birthweight_df, aes(x = fitted_1, y = resid_1)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals")

set.seed(123)
cv_splits <- vfold_cv(birthweight_df, v = 5)

models <- map(cv_splits$splits, ~ lm(bwt ~ gaweeks + delwt + babysex, data = analysis(.x)))

cv_rmse_results <- map_dbl(models, ~ {
  preds <- predict(.x, newdata = analysis(cv_splits$splits[[1]]))
  sqrt(mean((preds - analysis(cv_splits$splits[[1]])$bwt)^2))
})

model_2 <- lm(bwt ~ blength + gaweeks, data = birthweight_df)

model_3 <- lm(bwt ~ bhead * blength * babysex, data = birthweight_df)

cv_results_2 <- map(cv_splits$splits, ~ lm(bwt ~ blength + gaweeks, data = analysis(.x)))
cv_results_3 <- map(cv_splits$splits, ~ lm(bwt ~ bhead * blength * babysex, data = analysis(.x)))

cv_rmse_results_2 <- map_dbl(cv_results_2, ~ {
  preds <- predict(.x, newdata = analysis(cv_splits$splits[[1]]))
  sqrt(mean((preds - analysis(cv_splits$splits[[1]])$bwt)^2))
})

cv_rmse_results_3 <- map_dbl(cv_results_3, ~ {
  preds <- predict(.x, newdata = analysis(cv_splits$splits[[1]]))
  sqrt(mean((preds - analysis(cv_splits$splits[[1]])$bwt)^2))
})

cv_rmse_results

cv_rmse_comparison <- tibble(
  Model = c("Model 1", "Model 2", "Model 3"),
  RMSE = c(mean(cv_rmse_results), mean(cv_rmse_results_2), mean(cv_rmse_results_3))
)

print(cv_rmse_comparison)
```